Size-environment interaction

### 2021-01-15 analysis/paper/vital rate models.Rmd
switched to using shts x ht for size

There are plants with 0 shoots and 0 ht (not sure what that means yet), so to log-transform, I need to first add a small number

```{r}
demog_full <-
  demog_full %>% 
  mutate(log_shts = log(shts + 0.0001), .after = shts) %>% 
  mutate(log_shts_next = log(shts_next + 0.0001), .after = shts_next) %>% 
  mutate(size = shts*ht, size_next = shts_next*ht_next)
```

@@

```{r}
g_post_cf <-
  gam(shts_next ~ 
        s(log_shts) + 
        s(plot, bs = "re") + #random effect of plot
        s(spei_history, L,
          bs = "cb",
          k = c(3, 35),  #unimodal-ish response to drought, but one knot per month lag.
          xt = list(bs = "cr")),
      family = poisson(link = "log"),
      # family = scat(link = "log"), #like leptokurtic gaussian.
      data = demog_post_cf,
      # gamma = 1.2, #I maybe need to optimize this with cross-validation?
      method = "REML")
```

@@

```{r}
appraise(g_post_cf)
```

Suggests possibly more knots for s(log_shts), but it looks like a straight line to me

@@

@@

```{r}
g_post_1ha <-
  gam(shts_next ~ 
        s(log_shts) +
        s(plot, bs = "re") +
        s(spei_history, L,
          bs = "cb",
          k = c(3, 35),  #unimodal response to drought, but one knot per month lag.
          xt = list(bs = "cr")),
      # family = poisson(link = "log"),
      # family = quasipoisson(link = "log"),
      family = gaussian(link = "log"),
      # family = scat(link = "log"), #like leptokurtic gaussian.
      data = demog_post_1ha,
      # gamma = 1.2, #increases smoothness when above 1
      method = "REML")
```

@@

```{r}
# use mean size plant across both habitats so plots are comparable
ref_data <-
  demog_post %>%
  dplyr::summarize(log_shts = log(mean(shts, na.rm = TRUE)),  #log of mean shoots, not mean of log_shts
                   plot = "newlevel")
# exp(1.032814) # 2.8 shoots
mean_size <- round(exp(ref_data$log_shts), 1)
```

@@

```{r}
plotdf <-
  pred_g_post_cf %>% 
  mutate(lag = round(lag)) %>% 
  group_by(x, lag) %>% 
  summarize(across(everything(),  mean))

@@

```{r}
s_post_cf <-
  gam(surv ~ 
        s(log_shts) +
        s(plot, bs = "re") +
        s(spei_history, L,
          bs = "cb",
          k = c(3, 35),  #cubic response (at most) to drought, one knot per month lag.
          xt = list(bs = "cr")),
      family = binomial,
      data = demog_post_cf,
      # gamma = 1.2, #increases smoothness when above 1
      method = "REML")
```
```{r}
# gam.check(s_post_cf) #not sure what to expect from any of these plots
appraise(s_post_cf)
```
```{r}
summary(s_post_cf)
```

```{r}
s_post_1ha <-
  gam(surv ~ 
        s(log_shts) +
        s(plot, bs = "re") +
        s(spei_history, L,
          bs = "cb",
          k = c(3, 35),  #cubic response (at most) to drought, one knot per month lag.
          xt = list(bs = "cr")),
      family = binomial,
      data = demog_post_1ha,
      # gamma = 1.2, #increases smoothness when above 1
      method = "REML")
```
```{r}
summary(s_post_1ha)
```

@@

@@

```{r}
f_post_cf <-
  gam(flwr ~ 
        s(log_shts) +
        s(plot, bs = "re") +
        s(spei_history, L,
          bs = "cb",
          k = c(3, 35),  #cubic response (at most) to drought, one knot per month lag.
          xt = list(bs = "cr")),
      family = binomial,
      data = demog_post_cf,
      # gamma = 1.2, #increases smoothness when above 1
      method = "REML")
```

@@

```{r}
f_post_1ha <-
  gam(flwr ~ 
        s(log_shts) +
        s(plot, bs = "re") +
        s(spei_history, L,
          bs = "cb",
          k = c(3, 35),  #cubic response (at most) to drought, one knot per month lag.
          xt = list(bs = "cr")),
      family = binomial,
      data = demog_post_1ha,
      # gamma = 1.2, #increases smoothness when above 1
      method = "REML")
```

@@

Drought also has a significant effect on survival in both habitat types.

>>>>

There are plants with 0 shoots and 0 ht.  I will remove those, as having no above ground biomass is really not a size, but a different category.
There's also a couple of plants that are probably data entry errors in height.

```{r}
demog_full <-
  demog_full %>% 
  filter(ht < 200, ht_next < 200) %>% 
  filter(shts > 0, shts_next > 0, ht > 0, ht_next > 0) %>% 
  mutate(size = shts*ht, size_next = shts_next*ht_next, log_size = log(size), log_size_next = log(size_next))
```

@@

```{r}
g_post_cf <-
  gam(log_size_next ~ 
        s(log_size) + 
        s(plot, bs = "re") + #random effect of plot
        s(spei_history, L,
          bs = "cb",
          k = c(3, 35),  #unimodal-ish response to drought, but one knot per month lag.
          xt = list(bs = "cr")),
      # family = gaussian(link = "identity"),
      family = scat(link = "identity"), #like leptokurtic gaussian.
      data = demog_post_cf,
      # gamma = 1.2, #I maybe need to optimize this with cross-validation?
      method = "REML")
```

@@

```{r}
AIC(g_post_cf)

appraise(g_post_cf)
```

AIC for... 
Gaussian: 40909.27. looks leptokurtic
scaled t: 37284.47.  Waaay better

@@

Hmm. Make sure this doesn't predict negative sizes.

@@

```{r}
g_post_1ha <-
  gam(log_size_next ~ 
        s(log_size) +
        s(plot, bs = "re") +
        s(spei_history, L,
          bs = "cb",
          k = c(3, 35),  #unimodal response to drought, but one knot per month lag.
          xt = list(bs = "cr")),
      family = scat(link = "identity"), #like leptokurtic gaussian.
      data = demog_post_1ha,
      # gamma = 1.2, #increases smoothness when above 1
      method = "REML")
```

@@

```{r}
# use mean size plant across both habitats so plots are comparable
ref_data <-
  demog_post %>%
  dplyr::summarize(log_size = log(mean(size, na.rm = TRUE)),  #log of mean size, not mean of log_size
                   plot = "newlevel")
# exp(1.032814) # 2.8 shoots
mean_size <- round(exp(ref_data$log_size), 1)
```

@@

```{r}
#TODO: replace with ridge plot
plotdf <-
  pred_g_post_cf %>% 
  mutate(lag = round(lag)) %>% 
  group_by(x, lag) %>% 
  summarize(across(everything(),  mean))

@@

```{r}
s_post_cf <-
  gam(surv ~ 
        s(log_size) +
        s(plot, bs = "re") +
        s(spei_history, L,
          bs = "cb",
          k = c(3, 35),  #cubic response (at most) to drought, one knot per month lag.
          xt = list(bs = "cr")),
      family = binomial,
      data = demog_post_cf,
      # gamma = 1.2, #increases smoothness when above 1
      method = "REML")
```
```{r}
# gam.check(s_post_cf) #not sure what to expect from any of these plots
appraise(s_post_cf, method = "simulate")
```
```{r}
summary(s_post_cf)
```
wow, not significant!

```{r}
s_post_1ha <-
  gam(surv ~ 
        s(log_size) +
        s(plot, bs = "re") +
        s(spei_history, L,
          bs = "cb",
          k = c(3, 35),  #cubic response (at most) to drought, one knot per month lag.
          xt = list(bs = "cr")),
      family = binomial,
      data = demog_post_1ha,
      # gamma = 1.2, #increases smoothness when above 1
      method = "REML")
```
```{r}
summary(s_post_1ha)
```
not significant!  I should try select=TRUE here.

@@

These are not significant, so probably don't need to actually make these plots.  I'm mostly just curious what a non-significant crossbasis function looks like.

@@

```{r}
f_post_cf <-
  gam(flwr ~ 
        s(log_size) +
        s(plot, bs = "re") +
        s(spei_history, L,
          bs = "cb",
          k = c(3, 35),  #cubic response (at most) to drought, one knot per month lag.
          xt = list(bs = "cr")),
      family = binomial,
      data = demog_post_cf,
      # gamma = 1.2, #increases smoothness when above 1
      method = "REML")
```

@@

```{r}
f_post_1ha <-
  gam(flwr ~ 
        s(log_size) +
        s(plot, bs = "re") +
        s(spei_history, L,
          bs = "cb",
          k = c(3, 35),  #cubic response (at most) to drought, one knot per month lag.
          xt = list(bs = "cr")),
      family = binomial,
      data = demog_post_1ha,
      # gamma = 1.2, #increases smoothness when above 1
      method = "REML")
```

@@

These aren't significant.

### 2021-01-15 analysis/supplementary-materials/size-environment_interaction.Rmd
switched to using shts x ht for size

@@

```{r data, echo=TRUE}
demog <-
  read_csv(
    here("analysis", "data", "derived_data", "ha_survey.csv"),
    col_names = TRUE,
    cols(plot = col_character(),
         bdffp_reserve_no = col_character())
  )
demog <-
  demog %>%
  mutate(log_shts = log(shts + 0.0001),
         log_shts_next = log(shts_next + 0.0001))
```

@@

```{r}
growth_y <- lmer(log_shts_next ~ log_shts + (1|plot) + (1|year), data = demog_post_cf)
growth_sy <- lmer(log_shts_next ~ log_shts + (1|plot) + (log_shts|year), data = demog_post_cf)
AICtab(growth_y, growth_sy)
```

```{r}
lrt_growth <- sy_lrt(growth_y, growth_sy, nsim = 500)
```
Upper = 4.539, observed = 45.689

@@

```{r}
surv_y <- glmer(surv ~ log_shts + (1|plot) + (1|year), family = binomial, data = demog_post_cf)
surv_sy <- glmer(surv ~ log_shts + (1|plot) + (log_shts|year), family = binomial, data = demog_post_cf)
AICtab(surv_y, surv_sy)
```

```{r}
lrt_surv <- sy_lrt(surv_y, surv_sy, nsim = 500)
```
Upper = 3.913, observed = 29.181

@@

```{r}
flwr_y <- glmer(flwr ~ log_shts + (1|plot) + (1|year), family = binomial, data = demog_post_cf)
flwr_sy <- glmer(flwr ~ log_shts + (1|plot) + (log_shts|year), family = binomial, data = demog_post_cf)
AICtab(flwr_y, flwr_sy)
```

```{r}
lrt_flwr <- sy_lrt(flwr_y, flwr_sy, nsim = 500)
```
upper = 4.439, observed = 1.77

@@

```{r}
growth_y_1ha <- lmer(log_shts_next ~ log_shts + (1|plot) + (1|year), data = demog_post_1ha)
growth_sy_1ha <- lmer(log_shts_next ~ log_shts + (1|plot) + (log_shts|year), data = demog_post_1ha)
AICtab(growth_y_1ha, growth_sy_1ha)
```

@@

```{r}
surv_y_1ha <- glmer(surv ~ log_shts + (1|plot) + (1|year), family = binomial, data = demog_post_1ha)
surv_sy_1ha <- glmer(surv ~ log_shts + (1|plot) + (log_shts|year), family = binomial, data = demog_post_1ha)
AICtab(surv_y_1ha, surv_sy_1ha)
```

@@

```{r}
flwr_y_1ha <- glmer(flwr ~ log_shts + (1|plot) + (1|year), family = binomial, data = demog_post_1ha)
flwr_sy_1ha <- glmer(flwr ~ log_shts + (1|plot) + (log_shts|year), family = binomial, data = demog_post_1ha)
AICtab(flwr_y_1ha, flwr_sy_1ha)
```

```{r}

lrt_flwr_1ha <- sy_lrt(flwr_y_1ha, flwr_sy_1ha, nsim = 500)

```

>>>>

# TODO

- ~~Test that random effect of year is significant for all models~~

@@

```{r data, echo=TRUE}
demog <-
  read_csv(
    here("analysis", "data", "derived_data", "ha_survey.csv"),
    col_names = TRUE,
    cols(plot = col_character(),
         bdffp_reserve_no = col_character())
  )
demog <-
  demog %>%
  filter(ht < 200, ht_next < 200) %>% 
  filter(shts > 0, shts_next > 0, ht > 0, ht_next > 0) %>% 
  mutate(size = shts*ht, size_next = shts_next*ht_next, log_size = log(size), log_size_next = log(size_next))
```

@@

```{r}
growth_null <- lmer(log_size_next ~ log_size + (1|plot), data = demog_post_cf)
growth_y <- lmer(log_size_next ~ log_size + (1|plot) + (1|year), data = demog_post_cf)
growth_sy <- lmer(log_size_next ~ log_size + (1|plot) + (log_size|year), data = demog_post_cf)
AICtab(growth_null, growth_y, growth_sy)
```
Significant interaction.

```{r}
#This takes a very long time and gives results qualitatively similar to comparing AIC
lrt_growth <- sy_lrt(growth_y, growth_sy, nsim = 500)
```

@@

```{r}
surv_null <- glmer(surv ~ log_size + (1|plot), family = binomial, data = demog_post_cf)
surv_y <- glmer(surv ~ log_size + (1|plot) + (1|year), family = binomial, data = demog_post_cf)
surv_sy <- glmer(surv ~ log_size + (1|plot) + (log_size|year), family = binomial, data = demog_post_cf)
AICtab(surv_null, surv_y, surv_sy)
```
no interaction

```{r}
lrt_surv <- sy_lrt(surv_y, surv_sy, nsim = 500)
```

@@

```{r}
flwr_null <- glmer(flwr ~ log_size + (1|plot), family = binomial, data = demog_post_cf)
flwr_y <- glmer(flwr ~ log_size + (1|plot) + (1|year), family = binomial, data = demog_post_cf)
flwr_sy <- glmer(flwr ~ log_size + (1|plot) + (log_size|year), family = binomial, data = demog_post_cf)
AICtab(flwr_null, flwr_y, flwr_sy)
```
No interaction

```{r}
lrt_flwr <- sy_lrt(flwr_y, flwr_sy, nsim = 500)
```

@@

```{r}
growth_null_1ha <- lmer(log_size_next ~ log_size + (1|plot), data = demog_post_1ha)
growth_y_1ha <- lmer(log_size_next ~ log_size + (1|plot) + (1|year), data = demog_post_1ha)
growth_sy_1ha <- lmer(log_size_next ~ log_size + (1|plot) + (log_size|year), data = demog_post_1ha)
AICtab(growth_null_1ha, growth_y_1ha, growth_sy_1ha)
```
significant interaction

@@

```{r}
surv_null_1ha <- glmer(surv ~ log_size + (1|plot), family = binomial, data = demog_post_1ha)
surv_y_1ha <- glmer(surv ~ log_size + (1|plot) + (1|year), family = binomial, data = demog_post_1ha)
surv_sy_1ha <- glmer(surv ~ log_size + (1|plot) + (log_size|year), family = binomial, data = demog_post_1ha)
AICtab(surv_null_1ha, surv_y_1ha, surv_sy_1ha)
```
No interaction or effect of year.

@@

```{r}
flwr_null_1ha <- glmer(flwr ~ log_size + (1|plot), family = binomial, data = demog_post_1ha)
flwr_y_1ha <- glmer(flwr ~ log_size + (1|plot) + (1|year), family = binomial, data = demog_post_1ha)
flwr_sy_1ha <- glmer(flwr ~ log_size + (1|plot) + (log_size|year), family = binomial, data = demog_post_1ha)
AICtab(flwr_null_1ha, flwr_y_1ha, flwr_sy_1ha)
```
significant interaction?

```{r}
lrt_flwr_1ha <- sy_lrt(flwr_y_1ha, flwr_sy_1ha, nsim = 500)
```
